---
title: "Gait_Analysis_1"
author: "Katelyn Vuong"
output: html_document
---

```{r}
#read data
rundata = read.table("HuGaDB_v2_various_09_16.txt", header = TRUE)
head(rundata, 3)
```

We will visualize the time series data of the RF and LF accelerometer and gyroscope time series data. 

```{r}
library(ggplot2)
library(gridExtra)
#PLOTS FOR GAIT DYNAMIC TIME SERIES

#choose a chunk of the data
start = 1
end = 500
sampledata = rundata[start:end,]

#plot right foot accelerator
color_XYZ = c("X" = "red", "Y" = "green", "Z" = "blue")
p_acc_rf = ggplot(sampledata, aes(x = start:end)) +
  geom_line(aes(y = acc_rf_x, color = "X")) +
  geom_line(aes(y = acc_rf_y, color = "Y")) +
  geom_line(aes(y = acc_rf_z, color = "Z")) +
  scale_color_manual("Direction", values = color_XYZ) + 
  labs(x = "Time",
       y = "Signal",
       title = "RF accelerator") +
  theme_bw()


#plot right foot gyroscope
p_gyro_rf = ggplot(sampledata, aes(x=start:end)) +
  geom_line(aes(y = gyro_rf_x, color = "X")) +
  geom_line(aes(y = gyro_rf_y, color = "Y")) +
  geom_line(aes(y = gyro_rf_z, color = "Z")) +
  scale_color_manual("Direction", values = color_XYZ) +
  labs(x= "Time", y= "Signal", title = "RF gyroscope") +
  theme_bw()

#plot left foot accelerator
p_acc_lf = ggplot(sampledata, aes(x=start:end)) +
  geom_line(aes(y = acc_lf_x, color = "X")) +
  geom_line(aes(y = acc_lf_y, color = "Y")) +
  geom_line(aes(y = acc_lf_z, color = "Z")) +
  scale_color_manual("Direction", values = color_XYZ) +
  labs(x= "Time", y= "Signal", title = "LF accelerator") +
  theme_bw()

#plot left foot gyroscope
p_gyro_lf = ggplot(sampledata, aes(x=start:end)) +
  geom_line(aes(y = gyro_lf_x, color = "X")) +
  geom_line(aes(y = gyro_lf_y, color = "Y")) +
  geom_line(aes(y = gyro_lf_z, color = "Z")) +
  scale_color_manual("Direction", values = color_XYZ) +
  labs(x= "Time", y= "Signal", title = "LF gyroscope") +
  theme_bw() 

grid.arrange(p_acc_rf, p_gyro_rf, p_acc_lf, p_gyro_lf, ncol = 1)
```

Now we will cluster the data points using hierarchical clustering and Ward's minimum variance method. We want to cluster the data based on the similarity of the time series of the RF acc in the X, Y, Z dimensions. 

```{r}
#HIERARCHICAL CLUSTERING

#HC on X, Y, Z time series from the RF accelerator data using Ward's minimum variance method 

hc_acc_rf_x = hclust(dist(sampledata$acc_rf_x), method = "ward.D2")
plot(hc_acc_rf_x)
rect.hclust(hc_acc_rf_x, k =3, border = 2:4)

hc_acc_rf_y = hclust(dist(sampledata$acc_rf_y), method = "ward.D2")
hc_acc_rf_z = hclust(dist(sampledata$acc_rf_z), method = "ward.D2")

#HC on the triplet time series (X,Y,Z) from the RF accelerator data
hc_acc_rf = hclust(dist(sampledata[,1:3]), method = "ward.D2")
plot(hc_acc_rf)
rect.hclust(hc_acc_rf, k =3, border = 2:4)

```
Now we will encode digital code sequences in the RF accelerometer in X, Y, Z dimensions. This gives us a set of visualizations that display how the state sequences evolve over time for each dimension. 

```{r}
library(dplyr)
#DIGITAL-CODING

#encoding the three separate series and the triplet time series
state_acc_rf = data.frame(x = cutree(hc_acc_rf_x, 3),
                          y = cutree(hc_acc_rf_y, 3),
                          z = cutree(hc_acc_rf_z, 3),
                          hc27 = cutree(hc_acc_rf, 27),
                          hc10 = cutree(hc_acc_rf, 10))
#naturally combine the three digital code sequences from X, Y, Z
nc = apply(state_acc_rf,1,  function(x) paste(x[1], x[2], x[3], sep=","))
state_acc_rf$nc = as.numeric(factor(nc, labels = 1:n_distinct(nc)))
head(state_acc_rf)
```

```{r}
#RESULTANT CODE SEQUENCES

#3-state code sequence for X,Y,Z RF accelerator time series
p_st_x = ggplot(state_acc_rf, aes(x = start:end, y=x, color = "X")) +
  geom_line() +
  scale_color_manual(values = color_XYZ)+
  labs(x = "Time", y="State", title = "RF acc X") +
       theme_classic() + 
       theme(legend.position = "none")

p_st_y = ggplot(state_acc_rf, aes(x = start:end, y=y, color = "Y")) +
  geom_line() +
  scale_color_manual(values = color_XYZ)+
  labs(x = "Time", y="State", title = "RF acc Y") +
       theme_classic() + 
       theme(legend.position = "none")

p_st_z = ggplot(state_acc_rf, aes(x = start:end, y=z, color = "Z")) +
  geom_line() +
  scale_color_manual(values = color_XYZ)+
  labs(x = "Time", y="State", title = "RF acc Z") +
       theme_classic() + 
       theme(legend.position = "none")

grid.arrange(p_st_x, p_st_y, p_st_z)
```

We will create additional plots to visualize the naturally combiined code sequence and two different clustered code sequences. This visualizes how the states evolve over time in different code sequences from the RF accelerometer data. 

```{r}
#naturally combined code sequence

p_nc = ggplot(state_acc_rf, aes(x = start:end, y = nc)) +
  geom_line(color = "orange") +
  labs( x = "Time", 
        y = "State",
        title = "Naturally-combinate (X,Y,Z)") +
  theme_classic()

#27-state code sequence for the triplet time series
p_hc27 = ggplot(state_acc_rf, aes(x = start:end, y = hc27)) +
  geom_line(color = "darkgreen") +
  labs( x = "Time", 
        y = "State",
        title = "Clustering-combinate (X,Y,Z), state: 27") +
  theme_classic()

p_hc10 = ggplot(state_acc_rf, aes(x = start:end, y = hc10)) +
  geom_line(color = "darkblue") +
  labs( x = "Time", 
        y = "State",
        title = "Clustering-combinate (X,Y,Z), state: 10") +
  theme_classic()

grid.arrange(p_nc, p_hc27, p_hc10)

```






